00-Data-Preprocessing

### 2021-05-20 r0density/raw-data-processing/00-Data-Preprocessing.Rmd
Recreated data package to include boundaries of the states

@@

Save data table:
```{r}
save(county_geo, file = "county_geo.RData", compress = "xz")
```

@@

<!--

Convert missing R's to zeros and calculate expansion variables:
```{r}
county_geo_clean <- county_geo %>%
  mutate(R = replace_na(R, 0),
         density_2 = density^2) %>%
  drop_na(commuters) %>%
  st_as_sf()
```

Create list of neighbors:
```{r}
county_nb <- poly2nb(county_geo_clean)
```

Check list of neighbors:
```{r}
summary(county_nb)
```

Notice that there are 9 counties that are isolates. Which counties are these?
```{r}
county_geo_clean[c(69, 546, 547, 549, 1226, 1876, 2974, 3146, 3200),]
```

Find nearest neighbors by distance:
```{r}
county_nb_dist <- knearneigh(county_geo_clean %>% 
                               st_centroid()) %>%
  knn2nb()
```

Add nearest neighbor to isolates:
```{r}
county_nb[[69]] <- county_nb_dist[[69]]
county_nb[[546]] <- county_nb_dist[[546]]
county_nb[[547]] <- county_nb_dist[[547]]
county_nb[[549]] <- county_nb_dist[[549]]
county_nb[[1226]] <- county_nb_dist[[1226]]
county_nb[[1876]] <- county_nb_dist[[1876]]
county_nb[[2974]] <- county_nb_dist[[2974]]
county_nb[[3146]] <- county_nb_dist[[3146]]
county_nb[[3200]] <- county_nb_dist[[3200]]
```

Convert nearest neighbors to listw:
```{r}
nb_B <- nb2listw(county_nb, 
                 style="B", 
                 zero.policy = TRUE)
nb_B$style
```

Convert listw to sparse matrix for use in spatial tobit:
```{r}
B <- as(nb_B, "CsparseMatrix")
all(B == t(B))
```

## Replicate Sy et al.

Fit (mixed) linear models as in Sy et al.:
```{r}
table1model1 <- lme(R ~ density_log , 
                    random = ~ 1| state, 
                    data = county_geo_clean %>%
                      filter(R > 0))
summary(table1model1)
intervals(table1model1)
```

Linear mixed models of R0 and density + % private transportation
```{r}
table1model2 <- lme(R ~ density_log + private , 
                    random = ~ 1| state, 
                    data = county_geo_clean %>%
                      filter(R > 0))
summary(table1model2)
intervals(table1model2)
```

Linear mixed models of R0 and density + % private transportation + median income
```{r}
table1model3 <- lme(R ~ density_log + private + hincome, 
                    random = ~ 1| state, 
                    data = county_geo_clean %>%
                      filter(R > 0))
summary(table1model3)
intervals(table1model3)
```

Linear mixed models of R0 and density + % private transportation + median income + interaction of density and % private transportation
```{r}
table1model4 <- lme(R ~ density_log*private + hincome, 
                    random = ~ 1| state, 
                    data = county_R0_demog)
summary(table1model4)
intervals(table1model4)
```

## Fit tobit version of models

```{r}
table1model1 <- censReg(R ~ density_log, 
                        left = 0,
                    data = county_geo_clean)
summary(table1model1)
```

Linear mixed models of R0 and density + % private transportation
```{r}
table1model2 <- censReg(R ~ density_log + private, 
                    data = county_geo_clean)
summary(table1model2)
```

Linear mixed models of R0 and density + % private transportation + median income
```{r}
table1model3 <- censReg(R ~ density_log + private + hincome, 
                    data = county_geo_clean)
summary(table1model3)
```

Tobit models of R0 and density + % private transportation + median income + interaction of density and % private transportation
```{r}
table1model4 <- censReg(R ~ density_log*private + hincome, 
                    data = county_geo_clean)
summary(table1model4)

## Spatially autoregressive tobit

Fit spatially autoregressive tobit:
```{r}
# Fit SAR Tobit
fit_sartobit <- sartobit(R ~ density_log + private + hincome_log,
                         B,
                         ndraw = 1000,
                         burn.in = 200, 
                         showProgress = TRUE,
                         data = county_geo_clean,
                         computeMarginalEffects = TRUE)
summary(fit_sartobit)
impacts(fit_sartobit)
```

With interactions:
```{r}
# Fit SAR Tobit
fit_sartobit <- sartobit(R ~ density_log + density_log:private + hincome_log,
                         B,
                         ndraw = 1000,
                         burn.in = 200, 
                         showProgress = TRUE,
                         data = county_geo_clean)
summary(fit_sartobit)
impacts(fit_sartobit)
```
FIGURE 2 - POPULATION DENSITY THRESHOLD

Compare density of counties with oubreaks vs counties without outbreak
```{r}
county_outbreak_yn %>%
  group_by(included) %>%
  summarise(mean = median(Density, na.rm=TRUE), 
            lower_IQR = summary(Density)[2],
            upper_IQR=summary(Density)[5], 
            min=summary(Density)[1], 
            max=summary(Density)[6])

summary(county_outbreak_yn$Density)
wilcox.test(county_outbreak_yn$Density~county_outbreak_yn$included)
```

Boxplot of counties with oubreaks vs counties without outbreak
```{r}
county_outbreak_yn$included <- as.factor(county_outbreak_yn$included)

ggplot(county_outbreak_yn, aes(included, Density)) + 
  geom_boxplot(fill="gray92", width=0.5) + 
  ylim(0,200) + 
  #theme_minimal() + 
  ylab("Population Density (population/sq km)") + 
  xlab("United States Counties") + 
  theme_minimal(base_size=13)
```

TABLE 1 - LINEAR MIXED MODELS

Linear mixed models of R0 and density:
```{r}
table1model1 <- lme(R ~ density_log , 
                    random = ~ 1| state, 
                    data = county_geo_clean %>%
                      filter(R > 0))
summary(table1model1)
intervals(table1model1)
```

Linear mixed models of R0 and density + % private transportation
```{r}
table1model2 <- lme(R ~ density_log + private , 
                    random = ~ 1| state, 
                    data = county_geo_clean %>%
                      filter(R > 0))
summary(table1model2)
intervals(table1model2)
```

Linear mixed models of R0 and density + % private transportation + median income
```{r}
table1model3 <- lme(R ~ density_log + private + hincome, 
                    random = ~ 1| state, 
                    data = county_geo_clean %>%
                      filter(R > 0))
summary(table1model3)
intervals(table1model3)
```

Linear mixed models of R0 and density + % private transportation + median income + interaction of density and % private transportation
```{r}
table1model4 <- lme(R ~ density_log*private + hincome, 
                    random = ~ 1| state, 
                    data = county_R0_demog)
summary(table1model4)
intervals(table1model4)
```

## TABLE 2 - SENSITIVITY ANALYSIS

a. Deaths only - Subset counties with R0 from deaths only

```{r}
county_R0_demog_a <- subset(county_R0_demog, !is.na(county_R0_demog$R_death))
```

Linear mixed models of R0 (death) and density 
```{r}
table2model1_a <- lme(R_death ~ density_log , random = ~ 1| state, data = county_R0_demog_a)
summary(table2model1_a)
intervals(table2model1_a)
```

Linear mixed models of R0 (death) and density + % private transportation
```{r}
table2model2_a <- lme(R_death~ density_log + private , random = ~ 1| state, data = county_R0_demog_a)
summary(table2model2_a)
intervals(table2model2_a)
```

Linear mixed models of R0 (death) and density + % private transportation + median income
```{r}
table2model3_a <- lme(R_death~ density_log + private + hincome, random = ~ 1| state, data = county_R0_demog_a)
summary(table2model3_a)
intervals(table2model3_a)
```

Linear mixed models of R0 (death) and density + % private transportation + median income + interaction of density and % private transportation
```{r}
table2model4_a <- lme(R_death~ density_log*private + hincome, random = ~ 1| state, data = county_R0_demog_a)
summary(table2model4_a)
intervals(table2model4_a)
```

b. No counties adjacent to high density counties - within 15 mile radius, high density > 144.0492 population/sq km
```{r}
county_R0_demog_b <- county_R0_demog[county_R0_demog$sensitivitydensity,]
```

Linear mixed models of R0 and density (no counties adjacent to high density counties)
```{r}
table2model1_b <- lme(R ~ density_log , random = ~ 1| state, data = county_R0_demog_b)
summary(table2model1_b)
intervals(table2model1_b)
```

Linear mixed models of R0 and density + % private transportation (no counties adjacent to high density counties)
```{r}
table2model2_b <- lme(R ~ density_log + private , random = ~ 1| state, data = county_R0_demog_b)
summary(table2model2_b)
intervals(table2model2_b)
```

Linear mixed models of R0 and density + % private transportation + median income (no counties adjacent to high density counties)
```{r}
table2model3_b <- lme(R ~ density_log + private + hincome, random = ~ 1| state, data = county_R0_demog_b)
summary(table2model3_b)
intervals(table2model3_b)
```

Linear mixed models of R0 and density + % private transportation + median income + interaction of density and % private transportation (no counties adjacent to high density counties)
```{r}
table2model4_b <- lme(R ~ density_log*private + hincome , random = ~ 1| state, data = county_R0_demog_b)
summary(table2model4_b)
intervals(table2model4_b)
```

c. Removing high influence counties - Cook's D over 4/N for each model)

Linear mixed models of R0 and density (no outliers)
```{r}
sample_size<-nobs(table1model1)
cooksd1 <- CookD(table1model1)
influential <- as.numeric(names(cooksd1)[(cooksd1 > (4/sample_size))])
county_R0_demog_c <- county_R0_demog[-influential, ]
table1model1_c <- lme(R ~ density_log , random = ~ 1| state, data = county_R0_demog_c)
summary(table1model1_c)
intervals(table1model1_c)
```

Linear mixed models of R0 and density + % private transportation (no outliers)
```{r}
sample_size<-nobs(table1model2)
cooksd1 <- CookD(table1model2)
influential <- as.numeric(names(cooksd1)[(cooksd1 > (4/sample_size))])
county_R0_demog_c <- county_R0_demog[-influential, ]
table1model2_c <- lme(R ~ density_log + private , random = ~ 1| state, data = county_R0_demog_c)
summary(table1model2_c)
intervals(table1model2_c)
```

Linear mixed models of R0 and density + % private transportation + median income (no outliers)
```{r}
sample_size<-nobs(table1model3)
cooksd1 <- CookD(table1model3)
influential <- as.numeric(names(cooksd1)[(cooksd1 > (4/sample_size))])
county_R0_demog_c <- county_R0_demog[-influential, ]
table1model3_c <- lme(R ~ density_log + private + hincome, random = ~ 1| state, data = county_R0_demog_c)
summary(table1model3_c)
intervals(table1model3_c)
```

Linear mixed models of R0 and density + % private transportation + median income + interaction of density and % private transportation (no outliers)
```{r}
sample_size<-nobs(table1model4)
cooksd1 <- CookD(table1model4)
influential <- as.numeric(names(cooksd1)[(cooksd1 > (4/sample_size))])
county_R0_demog_c <- county_R0_demog[-influential, ]
table1model4_c <- lme(R ~ density_log*private + hincome, random = ~ 1| state, data = county_R0_demog_c)
summary(table1model4_c)
intervals(table1model4_c)
```

>>>>

Retrieve the state boundaries for plotting:
```{r}
state_geo <- get_decennial(geography = "state", 
                            variables = "P001001", # Total population
                            year = 2010, 
                            geometry = TRUE, 
                            keep_geo_vars = TRUE) %>%
  rename(population = value) %>%
  select(-variable)
```

@@

Prepare state boundaries for saving
```{r}
state_geo <- state_geo %>%
  select(-CENSUSAREA, -starts_with("geometry"), geometry)
```

Save data table:
```{r}
save(county_geo, file = "county_geo.RData", compress = "xz")
save(state_geo, file = "state_geo.RData", compress = "xz")
```

@@
